---
title: "R code and example for Space-To-Event model with Bootstrap resampling"
author: "Arnaud Lyet and Thierry Chambert"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
  css: styles.css
  #word_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      tidy = TRUE)
```


```{r, echo = FALSE}
rm(list = ls(all = T))
```


### Load packages, source files and data:
```{r, include = TRUE}
# packages
library(devtools)

# Load source files, functions and data
source_url("https://raw.githubusercontent.com/arnaudlyet/Lyet_et_al_2022_space-to-event/main/S0_0_Initialize_Git.R")
load("BC_dataset.RData")
```


# Data prep, period adjustment and declaration of variables and parameters

### Check the data
```{r, echo = FALSE}
# Subset data for that species only
species <- unique(dat0$common_name)[1]
dat <- dat0
kable(head(dat)) # Check data
```



### Check daily activity pattern and visually select the peak activity period for analysis
```{r, echo = FALSE}
temp <-  data.frame(date.time = as.POSIXct(strptime(dat$time, "%H:%M:%S")))
ggplot(data = temp, aes(x = date.time)) + 
  geom_histogram(colour = "black", fill="gray90", binwidth = 3600, boundary = 0) +
  scale_x_datetime(date_labels = "%H:%M") +
  labs(title = NULL, x = "Time of Day", y = "Observations Count") +
  theme(legend.position = "none") +
  theme_classic() + theme(axis.text.x = element_text(angle=0), 
                          axis.title=element_text(size=rel(1.5)))
```

### Select appropriate period of peak activity for the analysis
```{r, include = TRUE}
# Pick the earliest and latest times
tstart_txt <- "03:00:00" 
tend_txt <- "23:59:59"
time_start <- parse_date_time(paste(tstart_txt, "UTC"), "HMS")
time_end <- parse_date_time(paste(tend_txt, "UTC"), "HMS")

# Pick times to remove possible inactivity period. Times chosen equals if no inactivity period. 
tbkstart_txt <- "12:00:00"  
tbkend_txt <- "19:00:00"
time_break_start <-  parse_date_time(paste(tbkstart_txt, "UTC"), "HMS")
time_break_end <- parse_date_time(paste(tbkend_txt, "UTC"), "HMS")

# Calculate total available time per day in min and seconds
period1 <- period_to_seconds(hms(tend_txt)) + 1 - period_to_seconds(hms(tstart_txt))
period2 <- period_to_seconds(hms(tbkend_txt)) - period_to_seconds(hms(tbkstart_txt))

tot.seconds <- (period1 - period2)
tot.minutes <- tot.seconds / 60 # total numbers of minutes

```


### Declare the parameters of the model
```{r, include = TRUE}
Area = 1 # area for the report of the density estimate (in square km)

Nboot = 50 # total number of bootstraps
det.period = 1 # duration of the window of detection (in second)
interval <- 10 # average time between two consecutive time-lapse (in minutes)

# number of daily time-lapses (detection events)
(n.timelapse <- floor(tot.minutes / interval))
```


### Summarizes the observations for each time period (here per second) 
```{r, include = TRUE}
period = paste(det.period, "sec") # in seconds
tmp <- as.POSIXct(dat$time, format = "%H:%M:%S", tz = "UTC") 

# group the observations by period of x seconds
dat$time <- strftime(floor_date(tmp, unit = period), format = "%H:%M:%S", tz = "UTC") 
dat <- dat %>% group_by(cam, common_name, date, time) %>%
  summarize(n = sum(n))
```



### Declare objects for detection matrix and to store estimates' results
```{r, include = TRUE}
# declare detection matrix for each possible time-lapses
detec <- matrix(NA, nrow = nrow(activ), ncol = ncol(activ) * n.timelapse,
    dimnames = list(rownames(activ), NULL))
# expand the activity/inactivity status of the camera to each time lapse
for (j in 1:ncol(activ)) {  
  idx = (j - 1) * n.timelapse + (1:n.timelapse)
  detec[, idx] <- activ[, j]
} # j

# declare object to store STE estimates
STE <- data.frame(Iter = 1:Nboot, NbDet = NA, EST = NA, SE = NA)
```



# Data analysis
```{r, echo = FALSE}
select <- function(x, index) x[index]
```


### Identify all possible HH:MM:SS during the day that fall within the peak activity period
```{r, include = TRUE}
all_times <- seq(time_start, time_end, by = period)
if (time_break_start != time_break_end) {
  all_times <- all_times[-which(all_times %in% seq(time_break_start, time_break_end, by = period))]
}
```


### The time-lapse and bootstrap resampling is as follow:
### (1) Sample n random daily time-lapses (n.timelapse) from all the available HH:MM:SS. ### (2) The same time-lapses are used for each day of the survey.
### (3) Operation 1 and 2 are repeated as many times as the number of bootstrap chosen (Nboot). 
```{r, include = TRUE}
sel_times <- replicate(n = Nboot, rnd.sample(avail = all_times, n_timelapse = n.timelapse))
```



### Loop over all k bootstrap iterations. Increment the progress in the console and update the detail text.
```{r, include = TRUE}
pb <- progress_estimated(Nboot)
k = 1
for (k in 1:Nboot) {

  # Increment the progress bar.
  purrr::map_int(Nboot, arduously_long_nchar, .pb = pb)

  # Go over the k sampled times (within 1 day) for "instant" (n seconds) snapshots
  photo_time <- sel_times[[k]]
  
  # Select only detection that occurred at photo times
  sel <- which(as.character(dat$time) %in% photo_time)
  
  # Reduce the dataset to only keep detection at photo time
  dat.sel <- unique(dat[sel, c("common_name", "date", "time", "cam", "n")])
  
  # Where and when detection happened
  detec <- detec
  for (i in 1:nrow(dat.sel)) {
    rw <- which(station == as.character(dat.sel$cam)[i])
    c1 <- which(survey_day == as.integer(ymd(dat.sel$date[i])))
    c2 <- which(photo_time == as.character(dat.sel$time[i]))
    cl <- ((c1 - 1) * n.timelapse) + c2
    detec[rw, cl] <- no_na(detec[rw, cl]) + 1
    rm(rw)
    rm(cl)
  } # i
  
  n_det <- sum(detec, na.rm = TRUE)
  STE[k, "NbDet"] <- n_det
  
  # Estimation of density
  # Remove Occasions (days) with no active camera
  nb_cam <- apply(detec, 2, len_no_na)
  photo = detec[, nb_cam > 0] # select columns with active cam
  
  # Redefine nb of active cameras per occasion, removing case of 0 active camera
  nb_cam <- apply(photo, 2, len_no_na) # recreate vector of active cams per occ
  
  # Find total area of active cameras at each occasion
  # Reset detection file
  detec[detec > 0] <- 0
  active.area.tmp <- (detec + 1) * cta[[1]]
  areaCam <- apply(active.area.tmp, 2, sum, na.rm = T)
  # should I remove the ones with no camera?
  
  ## Data
  S = NA
  # this line below resamples the camera active on this date
  for (p in 1:ncol(photo)) {
    S[p] <-
      get_Sja(rm_na(photo[, p])[sample(length(rm_na(photo[, p])))])
  }
  
  dat_ste <-
    list(toevent = matrix(S, ncol = ncol(photo)),
         censor = areaCam,
         A = Area)

  # Estimate abundance with Space-to-Event
  run = STE_estN_fn(dat_ste, -10, 10)
  EST_N = run$estN
  SE_N = run$SE_N
  
  # Save output
  fill_col <- which(is.na(STE[k, ]))
  STE[k, fill_col] <- c(EST_N, SE_N)

} # k
```


# Results

### Summarize results for STE
```{r, echo = FALSE}
nZeros <- sum(STE$NbDet == 0) / Nboot
STE_EST_N <- mean(STE$EST) / Area
STE_SE_N <- mean(STE$SE) / Area
STE_CI_N <- quantile(STE$EST, probs = c(0.025, 0.975)) / Area
avg_grp_size <- mean(dat[["n"]], na.rm = TRUE)
effort <- sum(activ)
nObs   <- nrow(dat0)
timeTotal <- tot.seconds

# Get bootstrap averaged results
resSTE <-
  data.frame(
    Species = species,
    N_timelapse = n.timelapse,
    N_Zeros = nZeros,
    Density = STE_EST_N,
    SE = STE_SE_N,
    Lower_95p_CI = STE_CI_N[1],
    Upper_95p_CI = STE_CI_N[2]
  )

```


### Table results STE
```{r, echo = FALSE}
kable(resSTE)
```